Graphical Tests for Power Comparison of Competing Designs

### 2012-03-27 experiment.tex
updated figures

The difficulty of the charts is largely determined by the sample size taken from the original data set. Samples of 24 \% generally result in higher accuracy as the extra information facilitates detection of a relationship. Figure \ref{accuracy_preds} shows the relationship between the number of correct responses relative to the total number of responses and how this relationship differs between Euclidian and polar coordinates. The bars in the Euclidian plot are overall a lighter color than they are in the polar plot. This indicates a higher level of accuracy at all sample sizes. For Euclidian coordinates, we can see a stair step pattern in the accuracy levels as sample size increases. Lower levels of accuracy, indicated by darker colors, become consistently less prevelant as sample size increases. The relationship between sample size and accuracy level is not as simple for polar coordinates. Sample sizes of 2\% and 4\% have very similar accuracy levels. After 6\% the stair step pattern becomes visible and similar to that of Euclidian coordinates. Sample size seems to have less of an effect on accuracy for polar coordinates than it does for Euclidian coordinates.

>>>>

Ideally, efficient detection of data characteristics will be possible at small sample sizes. The difficulty of the charts is largely determined by the sample size taken from the original data set. Samples of 24 \% generally result in higher accuracy as the extra information facilitates detection of a relationship. Figure \ref{accuracy_preds} shows the relationship between the number of correct responses relative to the total number of responses and how this relationship differs between Euclidian and polar coordinates. The bars in the Euclidian plot are overall a lighter color than they are in the polar plot. This indicates a higher level of accuracy at all sample sizes. For Euclidian coordinates, we can see a stair step pattern in the accuracy levels as sample size increases. Lower levels of accuracy, indicated by darker colors, become consistently less prevelant as sample size increases. The relationship between sample size and accuracy level is not as simple for polar coordinates. Sample sizes of 2\% and 4\% have very similar accuracy levels. After 6\% the stair step pattern becomes visible and similar to that of Euclidian coordinates. Sample size seems to have less of an effect on accuracy for polar coordinates than it does for Euclidian coordinates. Predicted values of the proportion of correct responses when explained by sample size, the type of chart, and the interaction between these two factors, show that there is a significant difference between the relationship between sample size and correct responses for polar and euclidian coordinates. Indeed, the proportion of correct responses increases more rapidly as sample size increases for Euclidian coordinates than it does for polar coordinates.

The effect of a reference line on accuracy and time was not significant for both types of charts.

### 2012-03-27 setup.tex
updated figures

Statistical tests of mean efficiency by wind direction are not particularly helpful in this situation:
while the difference in mean efficiency between the most efficient wind direction and the least efficient shows a statistically significant difference with a $p$-value $\le 10^{15}$ %2.2e-16$. 
However, out of the 34 other wind directions,  another 31 show significant differences in efficiency as well. This is much more a property of the large data size rather than practically usable differences.

@@

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
%   \centering
   \begin{tabular}{ccl}
   \includegraphics[width=0.375\linewidth]{Euclidian_NoLine.pdf} & \hspace{-.3in}
   \includegraphics[width=0.375\linewidth]{Euclidian_Line.pdf} &  \hspace{-.25in} \multirow{2}{*}[.55in]{  \includegraphics[width=0.2\linewidth]{legend.pdf}} \\
   \includegraphics[width=0.43\linewidth]{Polar_NoLine.pdf} &  \hspace{-.3in}
   \includegraphics[width=0.43\linewidth]{Polar_Line.pdf}  
   \end{tabular}
   \caption{All four competing designs: polar versus Euclidean charts in top and bottom row, with (right) and without (left) reference lines at 50\%. }
   \label{layouts}
\end{figure}

In spite of the contextual circular property of wind direction, the pattern in efficiency did not seem to stand out well during the exploration process, which led us to using a corresponding  Euclidean design. Both designs were additionally equipped with a reference line at 50\%. 
Figure \ref{layouts} shows an overview of all four chart types in the study.

@@

\begin{tabular}{cl}
\phantom{\includegraphics[width=0.03\linewidth]{Polar_Line.pdf}} & \vspace{-0.02in} \multirow{10}{*}{\hspace{-0.25in}\includegraphics[width=0.9\linewidth]{turk4-designs.pdf}} \\
\includegraphics[width=0.055\linewidth]{Polar_Line.pdf} \\
\includegraphics[width=0.055\linewidth]{Polar_NoLine.pdf} \\
\includegraphics[width=0.045\linewidth]{Euclidian_Line.pdf} \\
\includegraphics[width=0.045\linewidth]{Euclidian_NoLine.pdf}\\
\phantom{\includegraphics[width=0.03\linewidth]{Polar_Line.pdf}}\\
\includegraphics[width=0.055\linewidth]{Polar_Line.pdf} \\
\includegraphics[width=0.055\linewidth]{Polar_NoLine.pdf} \\
\includegraphics[width=0.045\linewidth]{Euclidian_Line.pdf} \\
      \includegraphics[width=0.045\linewidth]{Euclidian_NoLine.pdf}
  \end{tabular} 
  \vspace{0.1in}
   \caption{Power results for two competing designs: polar versus a Euclidean, with and without a reference line. Error bars indicate 95\% confidence intervals. Plots are facetted by percentage of data used. }
   \label{fig:example}
\end{figure}

>>>>

Statistical tests of mean efficiency by wind direction are not particularly helpful in this situation:
the difference in mean efficiency between the most efficient wind direction and the least efficient shows a statistically significant difference with a $p$-value $\le 10^{15}$. %2.2e-16$. 
However, out of the 34 other wind directions,  another 31 show significant differences in efficiency as well. This is much more a property of the large data size rather than practically usable differences. Mere significances also do not allow us an assessment of the underlying pattern.

@@

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
%   \centering
 \hspace{-.1in}
   \begin{tabular}{ccl}
   \includegraphics[width=0.375\linewidth]{euclid_noline.pdf} & \hspace{-.3in}
   \includegraphics[width=0.375\linewidth]{euclid_line.pdf} &  \hspace{-.2in} \multirow{2}{*}[.55in]{  \includegraphics[width=0.2\linewidth]{legend.pdf}} \\
   \includegraphics[width=0.43\linewidth]{Polar_NoLine.pdf} &  \hspace{-.3in}
   \includegraphics[width=0.43\linewidth]{Polar_Line.pdf}  
   \end{tabular}
   \caption{All four competing designs: polar versus Euclidean charts in top and bottom row, with (right) and without (left) reference lines at 50\%. }
   \label{layouts}
\end{figure}

In spite of the contextual circular property of wind direction, the pattern in efficiency did not seem to stand out well during the exploration process, which led us to using a corresponding  Euclidean design. Both designs were additionally equipped with a reference line at 50\%. 
Figure \ref{layouts} shows an overview of all four chart types in the study. During the exploration of the data, it became clear that Seattle airport functions at it best with winds coming from the East to Southeast, while winds from the West seemed to be most troublesome. Since runways can be used in both directions, changing the runway usage according to dominant wind direction for that day or time of the year might be a feasible solution in gaining efficiency.

@@

\begin{tabular}{cl}
\phantom{\includegraphics[width=0.03\linewidth]{Polar_Line.pdf}} & \vspace{-0.035in} \multirow{10}{*}{\hspace{-0.25in}\includegraphics[width=0.9\linewidth]{turk4-designs.pdf}} \\
\includegraphics[width=0.055\linewidth]{Polar_Line.pdf} \\
\includegraphics[width=0.055\linewidth]{Polar_NoLine.pdf} \\
\includegraphics[width=0.045\linewidth]{Euclidian_Line.pdf} \\
\includegraphics[width=0.045\linewidth]{Euclidian_NoLine.pdf}\\
\phantom{\includegraphics[width=0.03\linewidth]{Polar_Line.pdf}}\\
\includegraphics[width=0.055\linewidth]{Polar_Line.pdf} \\
\includegraphics[width=0.055\linewidth]{Polar_NoLine.pdf} \\
\includegraphics[width=0.045\linewidth]{Euclidian_Line.pdf} \\
      \includegraphics[width=0.045\linewidth]{Euclidian_NoLine.pdf}
  \end{tabular} 
  \vspace{0.1in}
   \caption{Power results for two competing designs: polar versus a Euclidean, with and without a reference line. Error bars indicate 95\% confidence intervals. Plots are facetted by percentage of data used. }
   \label{fig:example}
\end{figure}

An additional factor we were interested in this particular situation was to assess, how much of the data we actually needed to use in a design to have observers pick out the pattern. Clearly, a design is more efficient, if a smaller sample size is sufficient for showing the presence of a relationship. In order to investigate this, we took different size samples of the data and created lineup plots of all four designs for these subsets. Another perturbation is to 
That also allows us to get multiple responses from each observer to assess individual ability.
